---
title: "Data Visualization for Graduation Rate"
output:
  pdf_document: default
  html_document: default
---


```{r,include=FALSE}
library(ggplot2)
library(rvest)
library(dplyr)
library(scales)
library(reshape2)
load("~/Desktop/gr.Rdata")
states <- map_data("state")
```


Firstly, Let us take a look at the total number of students and graduation rate among the US.

```{r,echo=F}
g1 <- grnew %>%group_by(State) %>% summarise(count = sum(ALL_T,na.rm=T))
names(g1)[1] <- "region"
g1$region <- tolower(g1$region)
g1 <- merge(states,g1,by="region")
ggplot(data=g1,aes(x=long,y=lat,group=group,fill=g1$count))+geom_polygon(colour="white")+scale_fill_continuous(low = "#feceda", high = "#c81f49", guide="colorbar")+labs(fill = "Number of Students")+coord_map()+ggtitle("The number of students")


g2 <- grnew %>%group_by(State) %>% summarise(rate = sum(ALL_G,na.rm=T)/sum(ALL_T,na.rm=T))
names(g2)[1] <- "region"
g2$region <- tolower(g2$region)
g2 <- merge(states,g2,by="region")

ggplot(data=g2,aes(x=long,y=lat,group=group,fill=g2$rate))+geom_polygon(colour="white")+scale_fill_continuous(low = "white", high = "#c81f49", guide="colorbar")+labs(fill = " Graduation Rate")+coord_map()+ggtitle("The gradution rate of students")
```

We can see that California and Texas are two typical states with large amount of students and high graduation rate.

```{r,echo=F}
g5 <- grnew %>%group_by(State) %>% summarise(graduation=sum(ALL_G/100,na.rm=T),nongraduation=sum(ALL_T-ALL_G/100,na.rm=T))
g5 <- melt(g5,id="State")
ggplot(g5)+geom_bar(aes(x=State,y=value,fill=variable),position = "fill",stat="identity")+ggtitle("Proportion of graduation among states")
```

Then we take a look at what happens between different groups as well as the changes over years.

```{r,echo=F,warning=F}
g31 <- grnew%>%group_by(year)%>%summarise(total_rate=sum(ALL_G/100,na.rm=T)/sum(ALL_T,na.rm=T),type="total")
g32 <- grnew%>%group_by(year)%>%summarise(total_rate=sum(as.numeric(White_R)*as.numeric(White_T),na.rm=T)/sum(as.numeric(White_T)*100,na.rm=T),type="White")
g33 <- grnew%>%group_by(year)%>%summarise(total_rate=sum(as.numeric(Econ_R)*as.numeric(Econ_T),na.rm=T)/sum(as.numeric(Econ_T)*100,na.rm=T),type="Economically Disadvantaged")
g34 <- grnew%>%group_by(year)%>%summarise(total_rate=sum(as.numeric(LimitE_R)*as.numeric(LimitE_T),na.rm=T)/sum(as.numeric(LimitE_T)*100,na.rm=T),type="LimitEnglish")
g35 <- grnew%>%group_by(year)%>%summarise(total_rate=sum(as.numeric(Disa_R)*as.numeric(Disa_T),na.rm=T)/sum(as.numeric(Disa_T)*100,na.rm=T),type="Disability")
g36 <- grnew%>%group_by(year)%>%summarise(total_rate=sum(as.numeric(Hispanic_R)*as.numeric(Hispanic_T),na.rm=T)/sum(as.numeric(Hispanic_T)*100,na.rm=T),type="Hispanic")
g37 <- grnew%>%group_by(year)%>%summarise(total_rate=sum(as.numeric(AsianPasific_R)*as.numeric(AsianPasific_T),na.rm=T)/sum(as.numeric(AsianPasific_T)*100,na.rm=T),type="AsianPasific")
g38 <- grnew%>%group_by(year)%>%summarise(total_rate=sum(as.numeric(Black_R)*as.numeric(Black_T),na.rm=T)/sum(as.numeric(Black_T)*100,na.rm=T),type="Black")
g3 <- rbind(g31,g32,g33,g34,g35,g36,g37,g38)
ggplot(data=g3,aes(x=year,y=total_rate,col=type))+geom_line()+ggtitle("The graduation rate for different groups")

ggplot(grnew,aes(x=factor(year),y=ALL_R,fill=factor(year)))+geom_violin()+ggtitle("Distributions of overall graduation rate over years")

g4 <- grnew[,c(8,22,24,26,27)]
g4 <- melt(g4,id="year")
g4$value <- as.numeric(g4$value)
g4 <- g4[!is.na(g4$value),]

ggplot(g4,aes(x=value,col=factor(variable),fill=factor(variable)))+geom_density(alpha=0.06)+ggtitle("Distributions of graduation rate among groups")

```

It is obvious the graduation rate is incresing for all subgroups. In addition, the distribution of graduations of schools is moving to the left. Also, we can see a significant difference between overall rate and rate for subgroups.\\

Finally, we focus on the top school defined by more than 95% graduation rate and more than 100 graduation students.

```{r,echo=F}
topschool <- which(grnew$ALL_R>95&grnew$ALL_T>100)
grtop <- grnew[topschool,]
schname <- unique(grtop$SCHNAM)
grtop <- filter(grnew,SCHNAM %in% schname)
grtop <- grtop[!duplicated(grtop$SCHNAM),]
g6 <- arrange(grtop%>%group_by(State)%>%summarise(n=n()),desc(n))
other <- sum(g6$n[10:50])
g6 <- g6[1:10,]
g6[11,] <- c("OTHER",other)
ggplot(g6,aes(x="",y=as.numeric(n),fill=State))+geom_bar(stat="identity")+coord_polar("y",start=0)+ggtitle("Proprotion of top school among states")
```

10 states led by California and Texas have about 2/3 of top schools. We will only focus on top schools in Califonia. 

```{r,echo=F}
ca <- filter(grnew,grnew$ALL_R>95&grnew$ALL_T>100&State=="CALIFORNIA")
ca <- filter(ca,!is.na(Econ_R))
ggplot(ca,aes(x=Econ_R,y=ALL_R,col=factor(year)))+geom_point()+geom_jitter()+ggtitle("Relationship between overall graduation rate and economically disadvataged graduation rate")
```


Since we have longitudinal data, we may employ linear mixed effect model. With a large amount of variables compared with observations, a sparsity assumption can be made and L-1 penalization is added to the loss function to select variable. The model can be specified as following:
$$Y_i^T=\beta X_i+\alpha T+\epsilon_i$$
Where $\beta$ is fixed effect and $\alpha$ is random effect. In addition, we estimate the parameter by minimizing:
$$$$




