---
title: "Variable Selection in Longitudinal Data"
output: html_document
---


### Cross covariance function

We checked the state-level correlation of grants allocation and graduation rate.
In the state map, it shows that the majority states have a positive correlation
between their grants receiption and the graduation rate.

```{r,echo=F,warning=F}
library(data.table)
library(ggplot2)
ggdat1 <- fread("~/Desktop/ggdat1.csv",header=T)
ggplot() + 
  geom_polygon(data=ggdat1, aes(x=long, y=lat, group = group, fill=ggdat1$rho),colour="white") + 
  # scale_fill_continuous(low = "thistle2", high = "darkred", guide="colorbar") +
  scale_fill_continuous(low = 'light blue', high = 'pink', guide="colorbar") +
  theme_bw() + 
  labs(fill = "correlations" ,title = "Correlation Map between grants and graduation rate", x="", y="")
```


### Lasso in mixed model

Since we have longitudinal data, we may employ linear mixed effect model. With a large amount of variables compared with observations, a sparsity assumption can be made and L-1 penalization is added to the loss function to select variable. The model can be specified as following:
$$Y_i^T=\beta X_i+\alpha T+\epsilon_i$$
Where $\beta$ is fixed effect and $\alpha$ is random effect. In addition, we estimate the parameter by minimizing:
$$||Y-\beta X-\alpha T||_2^2+\lambda||\beta||_1$$

First we considered the ordinary lasso model, with cross validation to tune the penelization parameter.
```{r,echo=F,warning=F}
library(glmmLasso)
aggrDat <- fread("~/Desktop/aggrDat.csv",header=T)
# lm1 <- glmmLasso(fix = gr~`84.004`+`84.010`+`84.011`+`84.013`+`84.041`+`84.060`+`84.141`+`84.144`+`84.149`+`84.165`+`84.184`+`84.196`+`84.215`+`84.282`+`84.287`+`84.295`+`84.299`+`84.330`+`84.336`+`84.350`+`84.351`+`84.354`+`84.356`+`84.358`+`84.360`+`84.362`+`84.363`+`84.365`+`84.366`+`84.367`+`84.368`+`84.369`+`84.371`+`84.374`+`84.377`,rnd = list(region=~1+year), lambda = 2.2, data = aggrDat)
# summary(lm1)
library(glmnet)
library(boot)

lm2 <- cv.glmnet(x = as.matrix(aggrDat[,-c(1,2,38)]),y = as.matrix(aggrDat[,2]),family = 'gaussian',alpha = 1)
plot(lm2,main="LASSO with cross validation")

lm3 <- glmnet(x = as.matrix(aggrDat[,-c(1,2,38)]),y = as.matrix(aggrDat[,2]),family = 'gaussian',alpha = 1,lambda = c(exp(-6),1))
plot(lm3,main="Variable Selection in Naive LASSO")

ll <- seq(1,5,length.out = 1000)
res1 <- NULL
fits = list(NA)
for(i in 1:length(ll)){
  fiti <- glmmLasso(fix = gr~`84.004`+`84.010`+`84.011`+`84.013`+`84.041`+`84.060`+`84.141`+`84.144`+`84.149`+`84.165`+`84.184`+`84.196`+`84.215`+`84.282`+`84.287`+`84.295`+`84.299`+`84.330`+`84.336`+`84.350`+`84.351`+`84.354`+`84.356`+`84.358`+`84.360`+`84.362`+`84.363`+`84.365`+`84.366`+`84.367`+`84.368`+`84.369`+`84.371`+`84.374`+`84.377`,rnd = list(region=~1+year), lambda = ll[i], data = aggrDat)
  res1 <- rbind(res1,fiti$coefficients)
}
```


To further account for the random effect from time(year) on state-level graduation rate, we then consider the mixed-effect lasso model to choose the top 10 program grants that have the most significant coefficients deviate from 0.

```{r}
num = 0
i=1000
important1 <- NULL
while(num<11){
  a = which(res1[i,]!=0)
  if(length(important1)!=length(a)){
    add <- setdiff(names(a),important1)
    important1 <- c(important1,add)
    num <- length(important1)
  }
  i = i-1
}
important1[-1]

```

It shows that program "84.060","84.184","84,295","84.351","84.356","84.358","84.360","84.367" and "84.369" have more effect.
The top 10 program grants with significant impact on state-level graduation rate are:

84.060: Indian Education_Grants to Local Educational Agencies
84.295: Ready-To-Learn Television
84.351: Arts in Education
84.356: Alaska Native Educational Program
84.358: Rural Education
84.360: High School Graduation Initiative
84.369: Grants for State Assessments and Related Activities
84.367: Improving Teacher Quality State Grants
84.010: Title I Grants to Local Educational Agency
84.184: Safe and Drug-free Schools and Communities_National Programs

### Regression Tree with Random Effect

In addition, we also employ random forest to see which variable are more important.

```{r,echo=F,warning=F}
aggrDat_new <- aggrDat[,c(-1,-38)]
library(caret)
library(mlbench)
control <-trainControl( method="repeatedcv",number=10, repeats=3)
model <-train(gr~., data=aggrDat_new, method="rf", importance=TRUE, na.action=na.exclude, preProcess="scale", trControl=control)
importance <-varImp(model)
plot(importance,main="Importance of Feature")
```

It shows that program "84.060","84,295","84.358" and "84.360" have more effect.
We list the top 20 important program grants:

X84.060 (Indian Education_Grants to Local Educational Agencies)  100.00
X84.295 (Ready-To-Learn Television)  68.40
X84.358 (Rural Education)  65.66
X84.360 (High School Graduation Initiative)  64.64
X84.350 (Transition to Teaching)  54.54
X84.041  (Impact Aid) 51.20
X84.365 (English Language Acquisition State Grants)   48.39  
X84.141 (Migrant Education_High School Equivalency Program)  47.59
X84.369 (Grants for State Assessments and Related Activities)  46.63
X84.366  (Mathematics and Science Partnerships) 44.94
X84.011 (Migrant Education_State Grant Program)  43.66
X84.196 (Education for Homeless Children and Youth) 42.54
X84.336 (Teacher Quality Partnership Grants) 42.44
X84.330 (Advanced Placement Program (Advanced Placement Test Fee; Advanced Placement Incentive Program Grants))  41.42
X84.144 (Migrant Education_Coordination Program)  39.72
X84.149 (Migrant Education_College Assistance Migrant Program)  38.29
X84.377 (School Improvement Grants)  35.46
X84.013 (Title I State Agency Program for Neglected and Delinquent Children and Youth)  35.40
X84.165 (Magnet Schools Assistance)  35.18
X84.287 (Twenty-First Century Community Learning Centers)  34.40

As we can conclude from the random forest model and lass/Mixed-effect lasso model, there is some overlap of the "significant" grants on lifting the state-level graduation rates.

Some further analysis may include:
Case study for states that shows negative correlation on graduation rate and grants receipient;
Analysis on "important" program grants.
